Proposal: A Natural Language Programming Toolkit for Python

### 2001-05-11 technical/proposal/proposal.tex
minor edits; added a section on "Statistical Processing and Visualization"
svn/trunk@28

However, the students do not necessarily have any experience
programming or designing software systems.  Thus, when students
attempt to build NLP systems, they may spend a significant amount of
their time on programming tasks that are not directly related to the
course of the content.

@@

\item[Efficiency] The toolkit doe not need to be highly optimized for
runtime performance.  However, it should be efficient enough that
students can use their NLP systems to perform real tasks.

@@

Each module define a small set of interfaces, which give
specifications for the types of classes that can be implemented by the
module.  Each module also contains classes that implement each
interface.  Often, a module will contain more than one implementation
for the same interface.

@@

\vspace{1.2mm}\noindent
\begin{tabular}{||p{.21\textwidth}p{.7\textwidth}}
  \textbf{SimpleTokenType} & A \texttt{TokenType}s that is
  represented by a single Python \texttt{string}.  Examples of
  \texttt{SimpleTokenType} values are ``dog'' and ``walking.'' \\

@@

When learning to use the toolkit, students intuitions about each
entity in the toolkit will be guided by the entities names'.  It is
therefore important to use consistent, correct, and intuitive
nomenclature when naming the entities defined by the toolkit.
The documentation should be careful to always use correct
nomenclature; in particular, it should never conflate two concepts
with the same name.

@@

@@

Epydoc is currently under development.  Contact Edward Loper for more
information about Epydoc.  For examples of the HTML pages generated by
Epydoc, see the reference documentation for the toolkit's prototype
modules 
at ``\texttt{http://www.cis.upenn.edu/$\sim$edloper/nltk/nltk.html}''.

@@

@@

>>>>

However, the students do not necessarily have any experience
programming or designing software systems.  Thus, when students
attempt to build NLP systems, they may spend a significant amount of
their time on programming tasks that are not directly related to the
course content.

@@

\item[Efficiency] The toolkit does not need to be highly optimized for
runtime performance.  However, it should be efficient enough that
students can use their NLP systems to perform real tasks.

@@

Each module defines a small set of interfaces, which give
specifications for the types of classes that can be implemented by the
module.  Each module also contains classes that implement each
interface.  Often, a module will contain more than one implementation
for the same interface.

@@

\vspace{1.2mm}\noindent
\begin{tabular}{||p{.21\textwidth}p{.7\textwidth}}
  \textbf{SimpleTokenType} & A \texttt{TokenType} that is
  represented by a single Python \texttt{string}.  Examples of
  \texttt{SimpleTokenType} values are ``dog'' and ``walking.'' \\

@@

When learning to use the toolkit, students intuitions about each
entity in the toolkit will be guided by the entities names.  It is
therefore important to use consistent, correct, and intuitive
nomenclature when naming the entities defined by the toolkit.
The documentation should be careful to always use correct
nomenclature; in particular, it should never conflate two concepts
with the same name.

@@

\subsection{Statistical Processing and Visualization}

The Python standard library contains minimal support for statistical
processing and visualization.  A number of third-party Python
libraries provide basic statistical and visualization tools
\cite{numpy, statspy, scipy, vtk}.  These tools could be used, either
by the toolkit, or directly by students, to perform statistical
processing and to visualize the results.

However, most of these tools have limited capabilities, and may not be
flexible enough for use with the toolkit.  Another option is to use a
well-established system for statistical processing and visualization,
that can be accessed indirectly via Python.  One such tool is the R
language \cite{r}.  R is an open-source version of the S system for
statistical processing and visualization \cite{s}.  This system
provides the advanced statistical tools and well-designed
visualization systems.  The R language can be accesed from Python via
the R/S Python interface \cite{rspython}.

The decision of what third-party modules and tools to be used by the
toolkit will be made after a more thourough investigation of the
advantages and distadvantages of respective tools.  Furthermore,
third-party modules will only be incorperated into the toolkit once
the toolkit develops enough to warrent their use.  However, tools such
as R seem like very promising candidates for adding advanced features
to the toolkit.

@@

Epydoc is currently under development.  Contact Edward Loper for more
information about Epydoc.  For examples of the HTML pages generated by
Epydoc, see the reference documentation for the toolkit's prototype
modules at 
``\texttt{http://www.cis.upenn.edu/$\sim$edloper/nltk/prototype/nltk.html}''.

@@

\bibitem{numpy}{Numerical Python \\
    \texttt{http://numpy.sourceforge.net/}}

@@

\bibitem{r}{The R Project for Statistical Computing \\
    \texttt{http://www.r-project.org/}}

\bibitem{rspython}{The R/S-Plus Python Interface \\
    \texttt{http://www.omegahat.org/RSPython/}}

\bibitem{s}{The S System \\
    \texttt{http://cm.bell-labs.com/cm/ms/departments/sia/S/}}

\bibitem{scipy}{Scientific Python \\
    \texttt{http://starship.python.net/crew/hinsen/scientific.html}}

\bibitem{statspy}{stats.py: A Python Statistical Module \\
    \texttt{http://www.vex.net/parnassus/apyllo.py/684222876.912038798}}

\bibitem{vtk}{VTK: The Visualization Toolkit \\
    \texttt{http://www.kitware.com/vtk.html}}
